---
title: "Cohorts"
subtitle: "Trace Functions for Exposure for Different Cohorts in a Population"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cohort}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

To construct a trace function to models exposure in different cohorts in the same population as they age:

```
?make_F_a
```

*** 

To understand or compare models for malaria epidemiology, we will often find it useful to construct trace functions describing exposure --- either the daily entomological inoculation rate (EIR) or the force of infection (FoI) --- in cohorts as they age. 

Average exposure in the population is specified by a [**composed time series** function](https://dd-harp.github.io/ramp.trace/articles/TimeSeries.html). 
$$E(t) = \bar X \times S(t) \times T(t) \times K(t)$$ 

If we want to compare cohorts in the same population, then we must acknowledge that exposure differs by age.  To model exposure, we construct a function to model relative biting rates by age:
$$\omega(a)$$ 
To model exposure for a cohort born on day $d,$ we note that time and age are related by $a = t-d.$ Exposure with respect to age for that cohort is thus:
$$E(a) = \bar X \times \omega (a) \times S(t-d) \times T(t-d) \times K(t-d).$$

*** 

# Example

## Exposure in a Population

```{r}
library(ramp.trace)
```

To illustrate, we specify a seasonal pattern, a trend, and a shock: 

```{r}
Sp <- makepar_F_sin()
Tp <- makepar_F_spline(tt=365*c(0:5), yy=c(1,1,1.6,.3,.7,1))
Kp <- makepar_F_sharkbite(D=260, L=300)
```

```{r fig.width=6, fig.height=4}
#| echo: false
F_s <- make_F_t(Sp)
F_t <- make_F_t(Tp)
F_k <- make_F_t(Kp)
tt = seq(0, 5*365, by = 10)
plot(tt, F_s(tt), type ="l", ylab = expression(list(S(t), T(t), K(t))), xlab = "Time (in Days)")
lines(tt, F_t(tt), type ="l")
lines(tt, F_k(tt), type ="l")
```

Over time, exposure in the population looks like this: 

```{r fig.width=6, fig.height=4} 
#| echo: false
F <- make_ts_function(scale = 3/365,season_par=Sp, trend_par=Tp, shock_par=Kp)
plot(tt, F(tt), type ="l", ylab = "Average Exposure", xlab = "Time (in Days)")
```

Suppose that relative biting rate by age looks like this:

```{r fig.width=6, fig.height=4}
Sa <- makepar_F_type2()
F_a <- make_F_t(Sa)
aa <- 1:3650
plot(aa/365, F_a(aa), type = "l", xlab = "a - Age (in Years)", ylab = expression(omega(a)))
```

Over the first three years of life, the patterns of exposure are quite different. In particular, exposure for the cohort born a year after the start of the study (dark red, dashed line) peaks in the first year of life. 

```{r fig.width=6, fig.height=8}
par(mfrow = c(2,1))
Fa <- make_F_a(avg = 3/365, age_par=Sa, season_par=Sp, trend_par=Tp, shock_par=Kp)
aa <- 1:1095
plot(aa/365, Fa(aa), col = "darkblue", type ="l", ylab = "Exposure", xlab = "Age (in Years)")
lines(aa/365, Fa(aa, d=365), col = "darkred", lty=2)
lines(aa/365, Fa(aa, d=730), col = "green3", lty =3)

plot(aa/365, cumsum(Fa(aa)), col = "darkblue", type ="l", ylab = "Cumulative Exposure", xlab = "Age (in Years)")
lines(aa/365, cumsum(Fa(aa, d=365)), col = "darkred", lty=2)
lines(aa/365, cumsum(Fa(aa, d=730)), col = "green3", lty =3)
```
