---
title: "ramp.xds"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ramp.xds}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

First, we define `F_season` and change the mean value of `Lambda.` To set up a seasonality function, we pass a parameter set from `ramp.trace::makepar_F_sin`: 

```{r}
library(ramp.xds)
library(ramp.trace)
```

```{r}
Lo = list(
       Lambda=400, 
       season_par=makepar_F_sin()
      )
```

We use the first model as a template for the new model, but we assign the return value a new name `model` so that the original one still exists: 

```{r}
model <- xds_setup( Loptions = Lo)
model <- burnin(model)
```

Now, we solve to get the orbits every five days over a three-year period.

```{r}
model <- xds_solve(model, Tmax=365*3, dt=5) 
```

After solving, we can plot the orbits. 

```{r, fig.height=2.6, fig.width=6, fig.cap = "**Figure 2:** Outputs with Seasonal Forced Emergence using a Trace Function" }
par(mfrow = c(1,2))
xds_plot_X(model)
xds_plot_M(model)
xds_plot_Y(model, add=T)
xds_plot_Z(model, add=T)
```
